---
title: "SF_ATBD"
author: "M Johnson mjohnson@bmrs.ie"
header-includes:
  - \usepackage{amsmath}

output: 
  bookdown::pdf_document2:
    number_sections: true 
bibliography: bib.bib
csl: citestyle.csl
---

```{r setup, include=FALSE, echo=FALSE}

options(scipen=999)

knitr::opts_knit$set(root.dir = '../')
library(ggplot2)
library(patchwork)

```

```{r refrun_Medulis, echo=FALSE, fig.cap='Reference runs of the model for M edulis, showing a) Chlorophyll concentration within the farm (ug/l) b) POC concentration in the farm (ug/l), c) Individual shell length (cm) ,d) daily NH4 production accross the farm (1km*1km),  e) soft tissue biomass (dry weight) per unit area (kg/m^2) and f) Fresh weight per unit area (kg/m^2)'}
source("run_SF.R")

HL<-reference_run(hclf,c(PHYC_data=FALSE))
LL<-reference_run(lclf,c(PHYC_data=FALSE))
LH<-reference_run(lchf,c(PHYC_data=FALSE))
HH<-reference_run(hchf,c(PHYC_data=FALSE))

  HL$run<-'hclf'
  LL$run<-'lclf'
  LH$run<-'lchf'
  HH$run<-'hchf'
  
  refruns<-rbind(HL,LL,LH,HH)

a<- ggplot(refruns) +
  aes(x = time, y = CHL_farm, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("CHL / ug.l^-1")+
  theme_minimal()
 b<-  ggplot(refruns) +
  aes(x = time, y = POC_farm, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("POC / ug.l^-1")+
  theme_minimal()
 c<-  ggplot(refruns) +
  aes(x = time, y = SHL, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("Shell length / cm")+
  theme_minimal()
  d<-   ggplot(refruns) +
  aes(x = time, y = NH4_production, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("NH4 yield / g.d^-1 ")+
  theme_minimal()

 e<-  ggplot(refruns) +
  aes(x = time, y = DSTW_PUA, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("Dry soft tissue / kg.m^-2")+
  theme_minimal()
  f<-   ggplot(refruns) +
  aes(x = time, y = FW_PUA, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("Fresh weight / kg.m^-2")+
  theme_minimal()

 (a|b)/(c|d)/(e|f)





 
 
 
```

```{r refrun_Cgigas, echo=FALSE, fig.cap='Reference runs of the model for C gigas, showing a) Chlorophyll concentration within the farm (ug/l) b) POC concentration in the farm (ug/l), c) Individual shell length (cm) ,d) daily NH4 production accross the farm (1km*1km),  e) soft tissue biomass (dry weight) per unit area (kg/m^2) and f) Fresh weight per unit area (kg/m^2)'}
source("run_SF.R")

HL<-reference_run(hclf,c(PHYC_data=FALSE),species='Cgig')
LL<-reference_run(lclf,c(PHYC_data=FALSE),species='Cgig')
LH<-reference_run(lchf,c(PHYC_data=FALSE),species='Cgig')
HH<-reference_run(hchf,c(PHYC_data=FALSE),species='Cgig')

  HL$run<-'hclf'
  LL$run<-'lclf'
  LH$run<-'lchf'
  HH$run<-'hchf'
  
  refruns<-rbind(HL,LL,LH,HH)

a<- ggplot(refruns) +
  aes(x = time, y = CHL_farm, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("CHL / ug.l^-1")+
  theme_minimal()
 b<-  ggplot(refruns) +
  aes(x = time, y = POC_farm, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("POC / ug.l^-1")+
  theme_minimal()
 c<-  ggplot(refruns) +
  aes(x = time, y = SHL, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("Shell length / cm")+
  theme_minimal()
  d<-   ggplot(refruns) +
  aes(x = time, y = NH4_production, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("NH4 yield / g.d^-1 ")+
  theme_minimal()

 e<-  ggplot(refruns) +
  aes(x = time, y = DSTW_PUA, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("Dry soft tissue / kg.m^-2")+
  theme_minimal()
  f<-   ggplot(refruns) +
  aes(x = time, y = FW_PUA, colour = run) +
  geom_point(shape = "circle", size = 0.7) +
  scale_color_hue(direction = 1) + ylab("Fresh weight / kg.m^-2")+
  theme_minimal()

 (a|b)/(c|d)/(e|f)





 
 
 
```
## Protein and energy content of shellfish species

### Protein content

Protein content as fraction of dry soft tissue weight. 

Mussel (M. edulis): 0.2-0.3 (assume average 0.25) [@Pleissner2012]
Oyster (C. Gigas): 0.5-0.57 (commercial population 0.53) [@Zhu2018]

Scallop P. Maximus @Cortes2021 19% by cooked weight assume no loss of moisture on cooking so assume 19% protein in drained wet tissue weight. DWW:DSTW = `r 1+allparams$species$defaults$P_maximus$parameters$WCT` so protein content = 0.19 * `r 1+allparams$species$defaults$P_maximus$parameters$WCT` = `r 0.14*(1+allparams$species$defaults$P_maximus$parameters$WCT)`


### Energy content

Energy content per individual is directly modelled by the DEB model. Energy content is therefore acessible in the model output - scaled to total farm output and per m2 as per other output variables. 
